\documentclass{article}


\title{Programming Statistical Models with Zelig 4}
\author{Matt Owen}


\usepackage{hyperref}


\begin{document}


\maketitle


% Introduction
\section{Introduction}
\label{intro}

In this course, students will develop statistical packages in the R programming
language using the Zelig 4 software suite and API. Emphasis is placed on using
generalized modeling techniques - linear regression, simulation, etc. - to
create predictive models that work with a large array of data-sets. Students will also
learn the basics of developing R packages for distribution via CRAN
(The Comprehensive R Network), a platform for sharing open-source statistical
software on the internet.

Registration can be found on the
\href{http://www.iq.harvard.edu/events/node/2710}{event site}
(\url{http://www.iq.harvard.edu/events/node/2710}).



% Prequisites
\section{Prerequisites}
\label{prereq}

Due to the brevity of the course, several requirements are imposed on registered
students:

\begin{itemize}
  \item {\bf Basic Programming}: conditional statements, loops and functions
  \item {\bf R Programming}: Sampling distributions and Linear Regressions
  \item {\bf Statistics}: Descriptive Statistics, linear regressions and
    bootstrapping
  \item {\bf Math}: Matrix Algebra
\end{itemize}



% Goals
\section{Goals}
\label{goals}

Upon completion of this course, students should have an understanding of the processes
involved in:

\begin{itemize}
  \item Developing statistical packages in R
  \item Programming Statistical and Mathematical ideas using Zelig
  \item Submitting packages to CRAN
\end{itemize}



% Goals
\section{Project}
\label{project}

In this course, students will choose from a list of several statistical models
to develop using the Zelig API. Workshops will be devoted to programming these
statistical models in R, with particular attention being paid to the process
of converting Mathematical ideas into functioning software. Additionally, 
students will learn practical approaches to programming and the design process
associated with developing modular software.



% Schedule
\section{Schedule}
\label{schedule}

This course will be divided between 4 days, each with separate lecture and
workshop component. Lectures will focus on introducing new concept and
programming techniques. Workshops will focus on the pragmatically implementing
the concepts and techniques introduced during the lecture. During the workshop,
emphasis will be placed on the problem-solving required to translate a
Mathematical idea into a full-featured R program.



% January 17th
\subsection{January 17\textsuperscript{th} \\ Creating R packages and Fitting 
Statistical Models}
  
This session focuses on the creation of R packages, presents a brief overview of
statistical simulation as it relates to creating R packages, and explains the
technical aspects of using model-fitting functions within the Zelig API.
  
\begin{itemize}

  % Lecture
  \item {\bf Lecture}
    \begin{itemize}
      \item Overview of programming statistical simulation
      \item The {\tt zelig} function
      \item Interfacing Zelig with external statistical packages
      \item Model Description and Citing Your Model
    \end{itemize}
    
  % Workshop
  \item {\bf Workshop}
    \begin{itemize}
      \item Structuring a statistical package
      \item Choosing a statistical model to create
      \item Writing the {\tt zelig2} function
    \end{itemize}

\end{itemize}



% January 18th
\subsection{January 18\textsuperscript{th} \\ Working with Data and
Counterfactuals}

This session focuses on the role of data and parameter simulation in the overall
process of statistical simulation. Additionally, students will learn the
technical aspects of parameter simulation within the Zelig API.

\begin{itemize}

  % lecture
  \item {\bf Lecture}
    \begin{itemize}
      \item The {\tt setx} function
      \item Simulating parameters of an external model
      \item Programming statistical bootstrapping
    \end{itemize}

  % workshop
  \item {\bf Workshop}
    \begin{itemize}
      \item Writing the {\tt param} method
      \item Specifying Explanatory Variables and Counterfactuals within Zelig
      \item Simulating Parameters
    \end{itemize}

\end{itemize}



% January 19th
\subsection{January 19\textsuperscript{th} \\ Simulating Quantities of Interest}

This session focuses on the role of the actual simulation of \emph{quantities of interest}.
That is, through the combined work of {\bf fitting statistical models} and {\bf parameter
simulation}, we are able to produce values with significant and particular meaning.
Emphasis will be placed on developing both the qualitative and quantitative aspects of the
process as well as the general applicability of the procedure.

\begin{itemize}

  % lecture
  \item {\bf Lecture}
    \begin{itemize}
      \item Understanding Statistical Simulation
      \item The {\tt sim} function
      \item Simulating quantities of interest
    \end{itemize}
	
  % workshop
  \item {\bf Workshop}
    \begin{itemize}
      \item Writing the {\tt qi} method
      \item Making statistical predictions with Zelig
    \end{itemize}

\end{itemize}



% January 20th
\subsection{January 20\textsuperscript{th} \\ Overview and Problem-Solving Workshop}

This session is devoted to the final aspects of R package creation: solving problems, 
beautifying-output and submitting packages to CRAN. As a result, students will spend most of this
day completing their statistical package and working out solutions to the inevitable pitfalls
that accompany developing statistical packages.

\begin{itemize}

  % lecture
  \item {\bf Lecture}
    \begin{itemize}
      \item How to submit packages to CRAN
      \item Review of the complete process of developing statistical packages
      \item What's next?
    \end{itemize}
		
  % workshop
  \item {\bf Workshop}
    \begin{itemize}
      \item Problem-solving workshop
      \item Putting it all together
    \end{itemize}
	
\end{itemize}

\end{document}
