% @author Matt Owen
% @date   2/22/2012
% This document serves as a specification - and talking point - to clarifying
% and elaborating on the role, function and definition of model formulae in
% the Zelig software suite (version 4.0+)
\documentclass{article}

% Packages
\usepackage{hyperref}

\title{Zelig Formula Specification}
\author{Matt Owen}

\newcommand{\tweedly}[0]{$\sim${ }}

\begin{document}

\maketitle



%
%
%
\section{Introduction and Motivation}
\label{sec:intro}

The following is a technical specification for the types of model formula
objects that are, should be, and could be officially supported by Zelig Core.

The primary motivation for this document is to create a non-rigid guideline for
the types of model formulae that can be used in the Zelig software suite
\emph{without} the need to write additional support functions. That is, so long
as model formula follow the specification defined in this document, there should
be little need to create or modify any code aside from the model's {\tt zelig2},
{\tt param} and {\tt qi} methods.

It should be noted that any issue stemming from the use of exotic or unsupported
formulae can be resolved with use of {\bf hook} functions or overloading the
{\bf \tt setx} or {\bf \tt sim} methods. This is purely a consequence of the
unexpected results that originate from constructing design matrices from
non-standard formulae.



%
%
%
\section{Categories of Formulae}
\label{sec:req}

The current implementation of Zelig allows for the direct support of four
categories of models:

\begin{itemize}

  \item {\bf A single response term, specified within a single model equation.}
    This is the case where users are studying a single response with a single
    set of predictor terms. This is the most basic model formula.

  \item {\bf A single response term, specified across multiple model equations.}
    This is the case where users are studying a single response across multiple
    sets of predictor terms.

  \item {\bf Multiple response terms, specified within a single model equation.}
    This is the case where users are studying the effects of a single set of
    predictor terms on multiple response terms.

  \item {\bf Multiple response terms, specified within multiple model equations.}
    This is the extreme case where users are studying multiple predictor terms
    dependent on different sets of predictor terms. Additionally, this type of
    specification is the most general.

\end{itemize}



%
%
%
\section{Existing Conventions in Zelig}
\label{sec:existing-zelig}

The categories specified in section \ref{sec:req}, have implementations in
Zelig (versions 3.5 and 4.0). It should be noted that the {\tt list}-method
\footnote{See the ``Multiple Equations'' row in section
\ref{subsec:table-conventions}} supported by Zelig is the only non-conventional
formula specification in the Zelig software package at the moment.


\subsection{Table of Conventions}
\label{subsec:table-conventions}

% Table
{\noindent}\begin{tabular}{|l|l|l|}

  % Top Border
  \hline

  % Column Names (Row #1)
  & Single Response Term & Multiple Response Terms \\ \hline

  % Row #2
  Single Equation &
  {\tt y \tweedly a + b + c} & {\tt cbind(x, y) \tweedly a + b + c}
  \\ \hline

  % Row #3
  Multiple Equations & 
  {\tt list(x \tweedly a, x \tweedly b)} &
  {\tt list(x \tweedly a, y \tweedly b)} \\ \hline


\end{tabular}


\subsection{User-end Specification}
\label{subsec:user-spec-formula}

The four types of formulae in section \ref{subsec:table-conventions} are created
using three basic ingredients:

\begin{itemize}

  \item {\bf ``formula'' objects}, which are the building block of any model
    formula

  \item {\bf the ``cbind'' function}, which - by convention - is used in simple
    linear models to specify multiple response terms. This, in general, does not
    work for every statistical model.

  \item {\bf ``list'' objects}, which allow users to specify multiple
    model equations in an intuitive fashion. This method is primarily used for
    the bivariate, dichotomous regressions.

\end{itemize}


\pagebreak


%
%
%
\section{Technical Specification for Basic Formula}

The following provides a specification for simple model formulae that can be
used out-the-box with Zelig core and its add-on packages.

It is worth nothing that the vast majority of statistical models in R use the
\verb|formula| object to specify information about the relationship between
\emph{response terms} and their \emph{predictor terms}. While in some cases,
this type of model formula is used to specify multiple response terms, for the
most part it is used to specify model formulae containing a singular response
term.


\subsection{Left-Hand Side of a Simple Formula}
\label{subsec:lhs}

The left-hand side of model equations largely follow the common practives
across all statistical models. However, several Zelig-isms exist to support
a unified framework for statistical analysis.


\begin{enumerate}

  \item The left-hand side of the formula specifies the response terms.

  \item The left-hand side supports the following operators and functions:
    {\tt +}, {\tt *}, and {\tt cbind}.

  \item Additionally, \verb|cbind| can be used as the first operator on the
    left-hand-side of an equation to specify multiple response variables. For
    more information on this method see subsection \ref{subsec:multiple}.

  \item The formula must only specify a single left-hand-side and a single
    right-hand-side. That is, nested formula are considered illegal. For
    example, the formula \begin{verbatim}y ~ a ~ b\end{verbatim} is considered 
    illegal.

  \item Mathematical operators (\verb|+|, \verb|*|, \verb|-|, \verb|/|) are
    interpretted arithematically. In particular, \verb|*| differs in meaning
    between the right and left sides of model formula.\footnote{Note that 
    {\tt *} does perform multiplication when placed on the right-hand side of
    the equation; it specifies that an equation is being modeled with
    interaction terms.}

\end{enumerate}


\subsection{Right-Hand Side of a Simple Formula}
\label{subsec:rhs}

\begin{enumerate}

  \item The right-hand side of the formula specifies the predictor terms.

  \item In additional to mathematical functions and arithmetic operators, Zelig
    can use the \verb|tag| formula to constrain parameter estimates across
    multiple equations. For more information, see section \ref{subsec:multiple}.

\end{enumerate}


\subsection{Formulae with Multiple Equations}
\label{subsec:multiple}

\begin{enumerate}

  \item The formula may be a list of model formulae, if it is describing
    multiple outcome variables. In this situation, each formulae must follow
    all of the above restrictions. Additionally, each element of the list may
    only specify a single response term.

  \item The {\tt tag} function can be used with special meaning in several
    linear models. In particular, {\tt tag} ensures that independent variables
    have identical effects upon separate response variables. This function is
    \emph{not} available for all generalized linear models.

  \item \verb|cbind| can only be used as the first operation on the left-hand side
    of an equation. That is, the formula
    {\tt cbind(x, y) + cbind(a, b) \tweedly 1} is considered illegal, despite
    the additon operation evaluating the left-hand side to a matrix.

\end{enumerate}



%
%
%
\section{Existing Conventions Outside of Zelig}
\label{sec:existing-elsewhere}

Before continuing onto the Zelig spec
There are several existing conventions that specify model formula from packages
outside of Zelig.

\begin{itemize}

  \item Basic model formula, designated using the \tweedly operator, specifya single
    model equation with a single response variable. Example: 
    {\tt y \tweedly a + b}

  \item Model formula designated using the \verb+cbind+ function in conjunction
    with the \tweedly operator specify multiple response terms for a single
    equation or set of predictor terms. Example: 
    {\tt cbind(x, y) \tweedly a + b + c}

  \item The ``Formula'', object found at
    \url{http://cran.r-project.org/web/packages/Formula/}, encapsulates the
    features of a standard ``formula'' object while extending its functionality.
    In particular, ``Formula'' objects allow for the specification of multiple
    response terms and equations without using a ``list'' object or a call to
    ``cbind''. Examples:
    \begin{itemize}
      \item {\tt Formula(y \tweedly a + b)}
      \item {\tt Formula(x | y \tweedly a + b)}
      \item {\tt Formula(y \tweedly a | b)}
      \item {\tt Formula(x | y \tweedly a | b)}
   \end{itemize}

   More on the ``Formula'' object specification can be found here:\\
   \url{http://cran.r-project.org/web/packages/Formula/index.html}

\end{itemize}



\end{document}
