\section{{\tt ei.dynamic}: Quinn's Dynamic Ecological Inference
Model}\label{ei.dynamic}

Given contingency tables with observed marginals, ecological inference
({\sc ei}) models estimate each internal cell value for each table.
Quinn's dynamic {\sc ei} model estimates a dynamic Bayesian model for
$2 \times 2$ tables with temporal dependence across tables (units).
The model is implemented using a Markov Chain Monte Carlo algorithm
(via a combination of slice and Gibbs sampling).  
For a hierarchical Bayesian implementation of {\sc ei} see Quinn's
dynamic {\sc ei} model (\Sref{ei.hier}). For contingency tables larger than
2 rows by 2 columns, see R$\times$C {\sc ei} (\Sref{ei.RxC}). 

\subsubsection{Syntax}

\begin{verbatim}
> z.out <- zelig(cbind(t0, t1) ~ x0 + x1, N = NULL, 
                 model = "MCMCei.dynamic", data = mydata)
> x.out <- setx(z.out, fn = NULL, cond = TRUE)
> s.out <- sim(z.out, x = x.out)
\end{verbatim}

\subsubsection{Inputs}
\begin{itemize}
\item \texttt{t0}, {\tt t1}: numeric vectors (either counts or
proportions) containing the column margins of the units to be
analyzed.

\item \texttt{x0}, {\tt x1}: numeric vectors (either counts or
proportions) containing the row margins of the units to be
analyzed.

\item \texttt{N}: total counts in each contingency table (unit).  If
\texttt{t0},\texttt{t1}, \texttt{x0} and \texttt{x1} are proportions,
you must specify \texttt{N}.

\end{itemize}

\subsubsection{Additional Inputs}

In addition, {\tt zelig()} accepts the following additional inputs for
\texttt{ei.dynamic} to monitor the convergence of the Markov chain:

\begin{itemize}
\item \texttt{burnin}: number of the initial MCMC iterations to be 
 discarded (defaults to 5,000). 

\item \texttt{mcmc}: number of the MCMC iterations after burnin
(defaults to 50,000).

\item \texttt{thin}: thinning interval for the Markov chain. Only every 
 \texttt{thin}-th draw from the Markov chain is kept. The value of 
\texttt{mcmc} must be divisible by this value. The default value is 1.

\item \texttt{verbose}: defaults to {\tt FALSE}.  If \texttt{TRUE}, the progress 
 of the sampler (every $10\%$) is printed to the screen.

\item \texttt{seed}: seed for the random number generator. The default
is \texttt{NA} which corresponds to a random seed of 12345. 

\end{itemize}

\noindent The model also accepts the following additional arguments to
specify priors and other parameters:

\begin{itemize}
\item \texttt{W}: a $p \times p$ numeric matrix describing the
structure of the temporal dependence among elements of $\theta_0$ and
$\theta_1$. The default value is 0, which constructs a weight matrix
corresponding to random walk priors for $\theta_0$ and $\theta_1$
(assuming that the tables are equally spaced throughout time, and that
the elements of \texttt{t0},
\texttt{t1},\texttt{x0},\texttt{x1} are temporally ordered). 

\item \texttt{a0}: $a_{0}/2$ is the shape parameter for the Inverse Gamma
prior on $\sigma_{0}^{2}$. The default is 0.825.

\item \texttt{b0}: $b_{0}/2$ is the scale parameter for the Inverse Gamma
prior on $\sigma_{0}^{2}$. The default is 0.0105.

\item \texttt{a1}: $a_{1}/2$ is the shape parameter for the Inverse Gamma
prior on $\sigma_{1}^{2}$. The default is 0.825.

\item \texttt{b1}: $b_{1}/2$ is the scale parameter for the Inverse Gamma
prior on $\sigma_{1}^{2}$. The default is 0.0105.
\end{itemize}

\noindent Users may wish to refer to \texttt{help(MCMCdynamicEI)} for more options.

\input{coda_diag}

\subsubsection{Examples}

\begin{enumerate}
 \item Basic examples \\
 Attaching the example dataset:
\begin{Schunk}
\begin{Sinput}
RRR>  data(eidat)
RRR> 
\end{Sinput}
\end{Schunk}
Estimating the model using \texttt{ei.dynamic}:
\begin{Schunk}
\begin{Sinput}
RRR>  z.out <- zelig(cbind(t0, t1) ~ x0 + x1, model = "ei.dynamic", 
+                  data = eidat,  mcmc = 40000, thin = 10, burnin = 10000, 
+                  verbose = TRUE)
RRR>  summary(z.out)
\end{Sinput}
\end{Schunk}
Setting values for in-sample simulations given the marginal values 
of {\tt t0}, {\tt t1}, {\tt x0}, and {\tt x1}:
\begin{Schunk}
\begin{Sinput}
RRR>  x.out <- setx(z.out, fn = NULL, cond = TRUE)
\end{Sinput}
\end{Schunk}
In-sample simulations from the posterior distribution:
\begin{Schunk}
\begin{Sinput}
RRR> s.out <- sim(z.out, x = x.out)
\end{Sinput}
\end{Schunk}
Summarizing in-sample simulations at aggregate level 
weighted by the count in each unit:
\begin{Schunk}
\begin{Sinput}
RRR>  summary(s.out)
\end{Sinput}
\end{Schunk}
Summarizing in-sample simulations at unit level for the first 5 units:
\begin{Schunk}
\begin{Sinput}
RRR>  summary(s.out, subset = 1:5)
\end{Sinput}
\end{Schunk}
\end{enumerate}
\clearpage
\subsubsection{Model}
Consider the following $2 \times 2$ contingency table for the racial
voting example.  For each geographical unit $i = 1, \dots, p$, the
marginals $t_i^0$, $t_i^1$, $x_i^0$, and $x_i^1$ are known, and we
would like to estimate $n_i^{00}$, $n_i^{01}$, $n_i^{10}$, and $n_i^{11}$.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{l|cc|c}
      & No Vote  & Vote &         \\
      \hline
      Black & $n_i^{00}$  & $n_i^{01}$ & $x_i^0$   \\
      White & $n_i^{10}$  & $n_i^{11}$ & $x_i^1$ \\
      \hline
      & $t_i^0$ & $t_i^1$ & $N_i$         
    \end{tabular}
  \end{center}
\end{table}

\noindent The marginal values $x_{i}^0$, $x_{i}^1$, $t_i^0$, $t_i^1$ are
observed as either counts or fractions. If fractions, the counts can
be obtained by multiplying by the total counts per table $N_i =
n_i^{00} + n_i^{01} + n_i^{10} + n_i^{11}$, and rounding to the
nearest integer.  Although there are four internal cells, only two
unknowns are modeled since $n_i^{01} = x_i^0 - n_{i}^{00}$ and
$n_{i}^{11} = s_i^1 - n_{i}^{10}$. \\

The hierarchical Bayesian model for ecological inference in 
$2 \times 2$ is illustrated as following: 
\begin{itemize}
\item The \emph{stochastic component} of the model assumes that
\begin{eqnarray*}
n_{i}^{00}\mid x_i^0,\beta_i^b &\sim& \textrm{Binomial}\left(
x_i^0, \beta_i^b\right)  ,\\
n_{i}^{10} \mid x_i^1, \beta_i^w &\sim& \textrm{Binomial}\left(
x_i^1, \beta_i^w\right)  \\
\end{eqnarray*}
where $\beta_{i}^{b}$ is the fraction of the black voters who vote
and $\beta_i^w$ is the fraction of the white voters who vote. $\beta_i^b$ 
and $\beta_i^w$ as well as their aggregate summaries are the focus
of inference.

\item The \emph{systematic component} of the model is 
\begin{eqnarray*}
\beta_i^b &=& \frac{\exp\theta_i^0}{1 - \exp\theta_i^0} \\
\beta_i^w &=& \frac{\exp\theta_i^1}{1 - \exp\theta_i^1} 
\end{eqnarray*}
The logit transformations of $\beta^b_i$ and $\beta^w_i$, $\theta_{i}^0$, 
and $\theta_i^1$ now take value on the real line. (Future versions may allow
$\beta_{i}^{b}$ and $\beta_{i}^{w}$ to be functions of observed covariates.)

\item The \emph{priors} for $\theta_{i}^0$ and $\theta_{i}^1$
 are given by
\begin{eqnarray*}
\theta_{i}^{0}   \mid\sigma_{0}^{2}&\propto & \frac{1}{\sigma_0^{p}} \exp
\left( -\frac{1}{2\sigma_{0}^2}\theta_0' P \theta_0 \right) \\
\theta_{i}^{1}   \mid\sigma_{1}^{2}&\propto & \frac{1}{\sigma_1^{p}} \exp
\left( -\frac{1}{2\sigma_{1}^2}\theta_1' P \theta_1 \right) 
\end{eqnarray*}
where $P$ is a $p \times p$ matrix whose off diagonal elements
$P_{ts}$ ($t \ne s$) equal $-W_{ts}$ (the negative values of the corresponding
elements of the weight matrix $W$), and diagonal elements 
$P_{tt}=\sum_{s\ne t} W_{ts}$. Scale parameters $\sigma_{0}^2$ and
$\sigma_{1}^2$ have hyperprior distributions as given below.  

\item The \emph{hyperpriors} for $\sigma_{0}^{2}$ and $\sigma_{1}^{2}$ are
given by
\begin{eqnarray*}
\sigma_{0}^{2}  & \sim& \textrm{Inverse Gamma}\left(\frac{a_{0}}{2},\frac{b_{0}}{2}\right)  ,\\
\sigma_{1}^{2}  & \sim& \textrm{Inverse Gamma}\left(\frac{a_{1}}{2},\frac{b_{1}}{2}\right)  ,
\end{eqnarray*}
where $a_{0}/2$ and $a_{1}/2$ are the shape parameters of the 
(independent) Gamma distributions while $b_{0}/2$ and $b_{1}/2$
 are the scale parameters. \\

The default hyperpriors for $\mu_{0},$ $\mu_{1}$, $\sigma_{0}^{2},$
and $\sigma_{1}^{2}$ are chosen such that the prior distributions for
$\beta^b$ and $\beta^w$ are flat.

\end{itemize}

\subsubsection{Output Values}

The output of each Zelig command contains useful information which you may
view. For example, if you run:

\begin{verbatim}
> z.out <- (cbind(t0, t1) ~ x0 + x1, N = NULL, 
          model = "ei.dynamic", data = mydata)
\end{verbatim}
 
\noindent then you may examine the available information in 
\texttt{z.out} by using \texttt{names(z.out)}, see the draws from the
posterior distribution of the quantities of interest by using
\texttt{z.out\$coefficients}, and view a default summary of
information through \texttt{summary(z.out)}. Other elements available
through the \texttt{\$} operator are listed below.

\begin{itemize}
\item From the \texttt{zelig()} output object \texttt{z.out}, 
you may extract:

\begin{itemize}
\item \texttt{coefficients}: draws from the posterior distributions
of the parameters.
\item \texttt{data}: the name of the input data frame.
\item \texttt{N}: the total counts when the inputs are fractions.
\item \texttt{seed}: the random seed used in the model.
\end{itemize}

\item From \texttt{summary(z.out)}, you may extract:
\begin{itemize}
\item \texttt{summary}: a matrix containing the summary information of the 
posterior estimation of $\beta^b_i$ and$\beta^w_i$ for each unit and
the parameters $\mu_0$, $\mu_1$, $\sigma_1$ and $\sigma_2$ based on
the posterior distribution.  The first $p$ rows correspond to
$\beta_i^b$, $i=1,\ldots p$, the row names are in the form of
\texttt{p0table}$i$. The $(p+1)$-th to the $2p$-th rows correspond to
$\beta_i^w$, $i=1,\ldots,p$. The row names are in the form of
\texttt{p1table}$i$. The last four rows contain information about
$\mu_0$, $\mu_1$, $\sigma_0^2$ and $\sigma_1^2$, the prior means and
variances of $\theta_0$ and $\theta_1$. 
\end{itemize}
\item From the \texttt{sim()} output object \texttt{s.out}, you may
extract quantities of interest arranged as arrays indexed by
simulation $\times$ column $\times$ row $\times$ observation, where
column and row refer to the column dimension and the row dimension of
the ecological table, respectively. In this model, only $2 \times 2$
contingency tables are analyzed, hence column$=2$ and row$=2$ in all
cases. Available quantities are:
\begin{itemize}
\item \texttt{qi\$ev}: the simulated expected values of each internal
cell given the observed marginals.
\item \texttt{qi\$pr}: the simulated expected values of each internal
cell given the observed marginals.
\end{itemize}

\end{itemize}

\subsection*{How to Cite}
\input{cites/ei.dynamic}
\input{citeZelig}
\subsection*{See also}
\emph{ei.dynamic} function is part of the MCMCpack library by Andrew D. Martin and Kevin M. Quinn \citep{MarQui05}.
The convergence diagnostics are part of the CODA library by Martyn Plummer, Nicky Best, Kate Cowles, and Karen Vines \citep{PluBesCowVin05}. Sample date are adapted from \cite{MarQui05}.
